A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-14 docs/icccn2009/paper.tex
Revise DSP Implementation, Broker, matcher and OSGi
git-svn-id: svn://localhost/netbeams/trunk@387 559edde2-a4d4-44e0-96fc-4035e40d7396

\begin{code}
interface DSPComponent
{
   public String getComponentType();
   public void initComponent(...);
   public void startComponent();
   public void stopComponent();
   public void deliver(DSPMessage msg);
   // ...
}
\end{code}

By implementing the above Java interface, the resulting code
effectively becomes a DSP Component that can be deployed on a DSP at
runtime. A component is identified by its type (line 3). Only one
component of a certain type per DSP is permissible. A type identifier
must be unique across all DSP Components in a DSP.  Upon deploying a
component, an initialization function is called (line 4). Once a
component is initialized, it may be started (line 5) and stopped (line
6) multiple times. Whenever the DSP wishes to pass a message (that was
generated by some DP), it invokes the \texttt{deliver()} method (line
7). The message structure will be explained in the next section.
One important feature of the DSP is the ability to redeploy itself (maybe
with a new version with updates) at runtime. This allows DSPs to be updated
remotely. This process is managed through messages sent to a special
DSP Component that is able to manage its local DSP.

@@

\subsection{DSP Broker}
\label{SEC_DSP_BROKER}

In order to keep the API symmetrical, the DSP is itself implemented as
a DSP Component, the so-called DSP Broker. While technically the DSP
Broker has the same API as any component, it serves a special role and
cannot be removed. When a DP sends a message, it actually
invokes the \texttt{deliver()} method of the DSP Broker. The broker is
responsible for routing the message to one or several DCs. Furthermore, it
distinguishes between local and remote destinations:

\begin{itemize}
\item Local delivery of messages: it uses local object reference to
  deliver the DSP message;
\item Remote delivery of messages: it uses a DSP Component marked as a
  default gateway to handle remote message delivery.
\end{itemize}

The selection of the DCs for a given message to where the broker will
deliver messages to is determined by a matcher.  The matcher is
responsible to find the possible set of DCs for a given DSP message,
and it relies on matching rules defined statically or at runtime. In
this way, a set of matching rules is defined in XML and stored at the
DSP run-time directory, where each matching rule is uniquely
identified by a rule ID and determines the following (see Figure
\ref{FIG_MATCHER_RULE}):

@@

When analyzing the matching rules, the matcher only selects the rules
that satisfy the matching criteria and forwards them back to the
broker. Once the broker has the matching rules, it determines whether
the delivery is done directly to a component defined on the match
target, or if the message is delivered to the optional gateway. We
have defined special gateway DSP components that is used to transport
the DSP Messages, serialized in XML, to any remote DSP instance.

@@

As mentioned earlier, the DSP is implemented in Java using the OSGi
platform. The so-called bundles, which are self-contained Java JAR files, can
be deployed and undeployed at runtime via special purpose class loaders. In 
fact, the API of the DSP is inspired by OSGi. All DSP Components depicted in Figure
\ref{FIG_NETBEAMS} are implemented as OSGi bundles. The DSP Broker discussed in
Section \ref{SEC_DSP_BROKER} is mapped to an OSGi bundle as well.

>>>>

\begin{code}
interface DSPComponent
{
   public String getComponentType();
   public void initComponent(String     id, 
                             DSPContext context);
   public void startComponent();
   public void stopComponent();
   public void deliver(DSPMessage msg);
   // ...
}
\end{code}

By implementing the above Java interface, the resulting code
effectively becomes a DSP Component that can be deployed on a DSP at
runtime. A component is identified by its type (line 3). Only one
component of a certain type per DSP is permissible. A type identifier
must be unique across all DSP Components in a DSP.  Upon deploying a
component, an initialization function is called (line 4). Once a
component is initialized, it may be started (line 4) and stopped (line
7) multiple times. Whenever the DSP wishes to pass a message (that was
generated by some DP), it invokes the \texttt{deliver()} method (line
8). DP sends messages through \texttt{Context} implementation (provided 
by the DSP during DSP initialization)The message structure will be explained in the next section.
One important feature of the DSP is the ability to redeploy itself (maybe
with a new version with updates) at runtime. This allows DSPs to be updated
remotely. This process is managed through messages sent to a special
DSP Component that is able to manage its local DSP.

@@

\subsection{DSP Broker}
Broker is responsible for routing the message to one or several DCs. It
distinguishes between local and remote destinations:

\begin{itemize}
\item Local delivery of messages: it uses local object reference to
  deliver the DSP message.
\item Remote delivery of messages: it uses a DSP Component marked as a
  gateway (configured in matching rules) to handle remote message delivery.
\end{itemize}

Broker provides the infrastructure necessary to deliver messages concurrently and
efficiently. Itis  also responsible for setting any additional message header 
information necessary for proper routing. Broker does not dictates which DCs should
receive a message. It delegates this responsibility to the Matcher.

\subsection{Matcher}

Matcher is the component in DSP responsible for matching messages from a DP to one or more
DCs. The decision process is based on evaluation of rules (see Figure
\ref{FIG_MATCHER_RULE}). Rules can be statically or dynamically
configured.

@@

Matcher's main responsibility is to shield DPs from the responsibility
for specifying consumers of their messages. DPs may create messages without any pre-determinted
DCs. Once the message is sent through DSP, the matcher will determine which DCs should
receive the message. Note that, a DP can still specify a DC for its message. In this case, Matcher
includes this DC in list of DCs to receive the message. Matcher also plays an important role on
the dynamic aspect of the DSP. Suppose the communication between two DSPs becomes unreliable and
it necessary to persist messages before they are sent through the communication channel. It is possible to
achieve this goal by deploying a DSP component which implements persistence and modifying the matching
rules to divert messages to the new DSP Component and from the component to the original destination.

@@

The DSP is implemented in java and serves as a container environment where 
DSP Components execute. Due to its micro-kernel, modular and dynamic characteristics 
the current DSP implementation uses OSGi \cite{osgi} as a deployment and execution 
infrastucture. One important DSP requirement is to impose on DSP components minimal 
dependency on DSP itself and no depency on OSGi. To this extent, all DSP resources
accessible to DSP components are specified by \texttt{DSPContext} interface.
DSP also uses message to communicate with DSP Components. It has a special "adapter"
that acts like a DSPComponent. Therefore, configuring DSP is no different from configuring
a DSP component.

DSP compoments are deployed as OSGi bundles. OSGi defines bundles life-cycle and
runtime upgrade mechanism. DSP uses these capabilities to allow remote deployment 
(bundles with DSP Components are trasmitted as message to the DSP Component Adaptor) 
and activation/deactivation of DSP Components.
